---
title: "House Prices"
author: "Kenneth Osborne"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: hide

---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE)
# knitr::opts_chunk$set(cache=TRUE)
myWD='C:/Users/Duckie/Documents/myR/kaggle/house_prices/'
knitr::opts_chunk$set(root.dir = myWD)
```

# Introduction



## Load libraries and helper functions

```{r, message = FALSE}
library('tidyverse') # all around good package
library('data.table') # data manipulation
library('dummies') # data manipulation
library('stringr') # text manipulation
library('corrplot') # visualization
library('party') # modelling
library('caret') # modelling
library('xgboost') #modelling
```


## Load Data

Here's a brief version of what you'll find in the data description file.

* SalePrice - the property's sale price in dollars. This is the target variable that you're trying to predict.
* MSSubClass: The building class
* MSZoning: The general zoning classification
* LotFrontage: Linear feet of street connected to property
* LotArea: Lot size in square feet
* Street: Type of road access
* Alley: Type of alley access
* LotShape: General shape of property
* LandContour: Flatness of the property
* Utilities: Type of utilities available
* LotConfig: Lot configuration
* LandSlope: Slope of property
* Neighborhood: Physical locations within Ames city limits
* Condition1: Proximity to main road or railroad
* Condition2: Proximity to main road or railroad (if a second is present)
* BldgType: Type of dwelling
* HouseStyle: Style of dwelling
* OverallQual: Overall material and finish quality
* OverallCond: Overall condition rating
* YearBuilt: Original construction date
* YearRemodAdd: Remodel date
* RoofStyle: Type of roof
* RoofMatl: Roof material
* Exterior1st: Exterior covering on house
* Exterior2nd: Exterior covering on house (if more than one material)
* MasVnrType: Masonry veneer type
* MasVnrArea: Masonry veneer area in square feet
* ExterQual: Exterior material quality
* ExterCond: Present condition of the material on the exterior
* Foundation: Type of foundation
* BsmtQual: Height of the basement
* BsmtCond: General condition of the basement
* BsmtExposure: Walkout or garden level basement walls
* BsmtFinType1: Quality of basement finished area
* BsmtFinSF1: Type 1 finished square feet
* BsmtFinType2: Quality of second finished area (if present)
* BsmtFinSF2: Type 2 finished square feet
* BsmtUnfSF: Unfinished square feet of basement area
* TotalBsmtSF: Total square feet of basement area
* Heating: Type of heating
* HeatingQC: Heating quality and condition
* CentralAir: Central air conditioning
* Electrical: Electrical system
* 1stFlrSF: First Floor square feet
* 2ndFlrSF: Second floor square feet
* LowQualFinSF: Low quality finished square feet (all floors)
* GrLivArea: Above grade (ground) living area square feet
* BsmtFullBath: Basement full bathrooms
* BsmtHalfBath: Basement half bathrooms
* FullBath: Full bathrooms above grade
* HalfBath: Half baths above grade
* Bedroom: Number of bedrooms above basement level
* Kitchen: Number of kitchens
* KitchenQual: Kitchen quality
* TotRmsAbvGrd: Total rooms above grade (does not include bathrooms)
* Functional: Home functionality rating
* Fireplaces: Number of fireplaces
* FireplaceQu: Fireplace quality
* GarageType: Garage location
* GarageYrBlt: Year garage was built
* GarageFinish: Interior finish of the garage
* GarageCars: Size of garage in car capacity
* GarageArea: Size of garage in square feet
* GarageQual: Garage quality
* GarageCond: Garage condition
* PavedDrive: Paved driveway
* WoodDeckSF: Wood deck area in square feet
* OpenPorchSF: Open porch area in square feet
* EnclosedPorch: Enclosed porch area in square feet
* 3SsnPorch: Three season porch area in square feet
* ScreenPorch: Screen porch area in square feet
* PoolArea: Pool area in square feet
* PoolQC: Pool quality
* Fence: Fence quality
* MiscFeature: Miscellaneous feature not covered in other categories
* MiscVal: $Value of miscellaneous feature
* MoSold: Month Sold
* YrSold: Year Sold
* SaleType: Type of sale
* SaleCondition: Condition of sale

```{r, warning=FALSE, message=FALSE, results=FALSE}
setwd('input')

training <- read.csv("train.csv") %>% cbind(Data="Training")
testing  <- read.csv("test.csv") %>% cbind(Data="Testing")
sample_sub  <- read.csv("sample_submission.csv")

df <- rbind(training %>% select(-SalePrice),testing)
```

## File structure

We inspect some basics

```{r}
df %>% dim

facts<- sapply(df,is.factor)
nums<- sapply(df,is.numeric)

cat(sum(facts), "factor variables and ", sum(nums), "numeric variables")
```

Checks on single valued variables and duplicate columns.

```{r}
cat("Duplicate Column Count:", nrow(df) - nrow(unique(df)))

col_ct = sapply(df, function(x) length(unique(x)))
cat("; Constant feature count:", length(col_ct[col_ct==1]))
```

Missing values

```{r}
missing_values <- df %>% summarize_all(funs(sum(is.na(.))))

missing_values <- gather(missing_values, key="feature", value="num_missing") %>% 
    cbind(is_factor = facts)

missing_values %>% 
    filter(num_missing>4) %>% 
    ggplot(aes(x=reorder(feature,num_missing),y=num_missing, fill=is_factor)) +
    xlab("Feature Name") +
    ylab("Number of Values Missing") +
    geom_bar(stat="identity")+
    coord_flip()+theme_bw()
```

There are so many missing values that they have to be dealt with.

# Data Wrangling

## Fix NA values 


We create a useful function for automatically adding a level named "None" to in place of NA.

```{r}
add_na_lvl <- function(x,na_lvl = which(x %>% is.na)){
    levels(x) <- c(levels(x),"None")
    x[na_lvl] <- "None"
    return(x)
}
```

### The Pools

Add new PoolQC feature for places with no pool area.

```{r}
noPools<- which(df$PoolArea==0 & is.na(df$PoolQC))

df$PoolQC <- add_na_lvl(df$PoolQC,noPools)
```


### Misc Feature, Alley, Fence, Fire Place

We apply the function to the next biggest interlopers.

```{r}
cleanme <- c("MiscFeature", "Alley", "Fence", "FireplaceQu")
df[,cleanme] <- apply(df[,cleanme],2,add_na_lvl) %>% as.data.frame()
```

### Lot Frontage

Sometimes there's no numeric value. Maybe that's the same as 0?
```{r}
which(df$LotFrontage==0)
```

It seems like these might be houses at the end of a street.

```{r}
df$LotFrontage[is.na(df$LotFrontage)]=0
```

### Garages

Make factor for places with no garage

```{r}
noGarage <- which(df$GarageArea==0)

cleanme <- c("GarageType", "GarageCond", "GarageFinish", "GarageQual")
df[,cleanme] <- apply(df[,cleanme],2, function(x) {
    add_na_lvl(x,noGarage)
}) %>% as.data.frame

df$GarageYrBlt[noGarage] <- 0
```

### Basements

Get rid of NA values in places without a basement.

```{r}
cleanme = c("BsmtQual","BsmtCond","BsmtExposure","BsmtFinType1","BsmtFinType2")
noBsmnt <- which(df$BsmtFinSF1==0 & df$BsmtFinSF2==0 & df$BsmtUnfSF==0)

df[noBsmnt,cleanme] %>% as.tbl()

df[,cleanme] <- apply(df[,cleanme],2, function(x) {
    add_na_lvl(x,noBsmnt)
}) %>% as.data.frame
```

### Masonry

```{r}
df$MasVnrType <- df$MasVnrType %>% add_na_lvl
df$MasVnrArea[df$MasVnrArea %>% is.na()] <- 0
```

## Check NA vals

And then re-evaluate

```{r}
missing_values <- df %>% summarize_all(funs(sum(is.na(.))))

missing_values <- gather(missing_values, key="feature", value="num_missing") %>% 
    cbind(is_factor = facts)

missing_values %>% 
    filter(num_missing>0) %>% 
    ggplot(aes(x=reorder(feature,num_missing),y=num_missing, fill=is_factor)) +
    xlab("Feature Name") +
    ylab("Number of Values Missing") +
    geom_bar(stat="identity")+
    coord_flip()+theme_bw()
```

# Modelling

Looking to minimize the Root Mean Squared Logarithmic Error

```{r}
target <- training %>% mutate(SalePrice=log(SalePrice))
training <- df %>% filter(Data=="Training") %>% mutate(SalePrice = target$SalePrice)
testing <- df %>% filter(Data=="Testing")
```

make it all numbers

```{r, warning=FALSE}
numbr_train <- training
# numbr_train <- apply(training,2,as.numeric) %>% as.data.frame()
for (i in 1:dim(numbr_train)[2]) {
    numbr_train[,i] <- numbr_train[,i] %>% as.numeric()
}

numbr_test <- testing
# numbr_test <- apply(testing,2,as.numeric) %>% as.data.frame()
for (i in 1:dim(numbr_test)[2]) {
    numbr_test[,i] <- numbr_test[,i] %>% as.numeric()
}
```

Next we partition

```{r}
set.seed(4321)
trainIndex <- createDataPartition(numbr_train$SalePrice, p = 0.8, list = FALSE, times = 1)

n_training <- numbr_train[trainIndex,]
n_validation <- numbr_train[-trainIndex,]
```

Reformat for XGB

```{r}
foo <- n_training %>% select(-c(SalePrice,Id,Data))
bar <- n_validation %>% select(-c(SalePrice,Id,Data))
baz <- numbr_test %>% select(-c(Id,Data))

dtrain <- xgb.DMatrix(as.matrix(foo),label = n_training$SalePrice)
dvalid <- xgb.DMatrix(as.matrix(bar),label = n_validation$SalePrice)
dtest <- xgb.DMatrix(as.matrix(baz))
```

Set the parameters

```{r}
xgb_params <- list(colsample_bytree = 0.5, #variables per tree 
                   subsample = 1, #data subset per tree 
                   booster = "gbtree",
                   max_depth = 6, #tree levels
                   eta = 0.3, #shrinkage
                   eval_metric = "rmse", 
                   objective = "reg:linear",
                   seed = 4321
                   )

watchlist <- list(train=dtrain, valid=dvalid)
```

Run the thing

```{r}
set.seed(4321)
gb_dt <- xgb.train(params = xgb_params,
                   data = dtrain,
                   print_every_n = 25,
                   watchlist = watchlist,
                   nrounds = 300)
```

```{r}
imp_matrix <- xgb.importance(feature_names = colnames(foo), model = gb_dt)

imp_matrix %>% filter(Gain>.005) %>% 
  ggplot(aes(reorder(Feature, Gain, FUN = max), Gain, fill = Feature)) +
  geom_col() +
  coord_flip() +
  theme(legend.position = "none") +
  labs(x = "Features", y = "Importance")
```

```{r}
pred1 <- predict(gb_dt,dtest)
pred1 <- exp(pred1)
```

Create output file

```{r}

pred <- testing %>% select(Id) %>% mutate(SalePrice=pred1)
pred %>% write_csv('submit.csv')
identical(dim(sample_sub),dim(pred))
```

```{r}
# set up the cross-validated hyper-parameter search
xgb_grid_1 = expand.grid(colsample_bytree = c(0.3,.04,0.5,0.6,0.7) #variables per tree 
                         subsample = 1, #data subset per tree 
                         booster = "gbtree",
                         max_depth = 6, #tree levels
                         eta = 0.3, #shrinkage
                         eval_metric = "rmse", 
                         objective = "reg:linear",
                         seed = 4321)
 
# pack the training control parameters
xgb_trcontrol_1 = trainControl(
method = "cv",
number = 5,
verboseIter = TRUE,
returnData = FALSE,
returnResamp = "all",                                                        # save losses across all models
classProbs = TRUE,                                                           # set to TRUE for AUC to be computed
summaryFunction = twoClassSummary,
allowParallel = TRUE
)
 
# train the model for each parameter combination in the grid,
#   using CV to evaluate
xgb_train_1 = train(
x = as.matrix(df_train %>%
select(-SeriousDlqin2yrs)),
y = as.factor(df_train$SeriousDlqin2yrs),
trControl = xgb_trcontrol_1,
tuneGrid = xgb_grid_1,
method = "xgbTree"
)
 
# scatter plot of the AUC against max_depth and eta
ggplot(xgb_train_1$results, aes(x = as.factor(eta), y = max_depth, size = ROC, color = ROC)) +
geom_point() +
theme_bw() +
scale_size_continuous(guide = "none")
```
